Data Analysis Methods for the Detection of the Epoch of Reionization

### 2017-11-20 PAPER.tex
-Read-through for typos/clarifcations (intro and sigloss toy model)

\author{
Carina Cheng\altaffilmark{1},
Aaron R. Parsons\altaffilmark{1,2},
Matthew Kolopanis \altaffilmark{3},
Daniel C. Jacobs\altaffilmark{3}, 
Saul A. Kohn\altaffilmark{4},
Adrian Liu\altaffilmark{1,5}, 
James E. Aguirre\altaffilmark{5},
Zaki S. Ali\altaffilmark{1}, 
Richard F. Bradley\altaffilmark{6,7,8},
Gianni Bernardi\altaffilmark{9,10,11}, 
Chris L. Carilli\altaffilmark{12,13},
David R. DeBoer\altaffilmark{2}, 
Matthew R. Dexter\altaffilmark{2},
Joshua S. Dillon\altaffilmark{1},
Pat Klima\altaffilmark{7},
David H. E. MacMahon\altaffilmark{2},
David F. Moore\altaffilmark{5},
Chuneeta D. Nunhokee\altaffilmark{14},
Jonathan C. Pober\altaffilmark{15}, 
William P. Walbrugh\altaffilmark{9},
Andre Walker\altaffilmark{9}
}

\altaffiltext{1}{Astronomy Dept., U. California, Berkeley, CA}
\altaffiltext{2}{Radio Astronomy Lab., U. California, Berkeley CA}
\altaffiltext{3}{School of Earth and Space Exploration, Arizona State U., Tempe AZ}
\altaffiltext{4}{Dept. of Physics and Astronomy, U. Penn., Philadelphia PA} 
\altaffiltext{5}{Berkeley Center for Cosmological Physics, Berkeley, CA} 
\altaffiltext{6}{Dept. of Electrical and Computer Engineering, U. Virginia, Charlottesville VA}
\altaffiltext{7}{National Radio Astronomy Obs., Charlottesville VA}
\altaffiltext{8}{Dept. of Astronomy, U. Virginia, Charlottesville VA}
\altaffiltext{9}{Square Kilometer Array, S. Africa, Cape Town South Africa}
\altaffiltext{10}{Dept. of Physics and Electronics, Rhodes University}
\altaffiltext{11}{Harvard-Smithsonian Cen. for Astrophysics, Cambridge MA}
\altaffiltext{12}{National Radio Astronomy Obs., Socorro NM}
\altaffiltext{13}{Cavendish Lab., Cambridge UK}
\altaffiltext{14}{Dept. of Physics and Electronics, Rhodes University, Grahamstown South Africa}
\altaffiltext{15}{Dept. of Physics, Brown University, Providence RI}

@@

\begin{abstract}
The Epoch of Reionization (EoR) is an uncharted era in our Universe's history during which the birth of the first stars and galaxies led to the ionization of neutral hydrogen in the intergalactic medium. There are many experiments investigating the EoR by tracing the $21$ cm line of neutral hydrogen, a signal which is very faint and difficult to isolate. With a new generation of instruments and a statistical power spectrum detection in our foreseeable future, it has become increasingly important to develop techniques that help maximize sensitivity and validate results. Additionally, it is imperative to understand the trade-offs between different methods and their effects on common power spectrum themes. In this paper, we focus on three major themes --- signal loss, power spectrum error bar estimation, and bias in measurements. We describe techniques that affect these themes using both a toy model and data taken by the 64-element configuration of the Donald C. Backer Precision Array for Probing the Epoch of Reionization (PAPER). In particular, we highlight how detailed investigations of these themes have led to a revised, higher $21$ cm power spectrum upper limit from PAPER-64. This revised result mostly stems from an updated signal loss calculation for loss associated with empirically estimated covariances, and we consider this limit to supersede previously published PAPER-64 and PAPER-32 results.
\end{abstract}

@@

This paper is organized into two main sections. In Section \ref{sec:Themes} we introduce the three themes of our focus, using a toy model to develop intuition for each one. In Section \ref{sec:CaseStudy} we present a case study into each theme using data from the PAPER-64 array, highlighting key changes from \citet{ali_et_al2015} which have led to a revised PAPER-64 power spectrum result (\cc{cite retraction or kolopanis paper}). We conclude in Section \ref{sec:Con}.

@@

The following techniques each have advantages when it comes to maximizing sensitivity and understanding systematics in data. However, some have limitations, and we will discuss circumstances in which there are trade-offs. We choose to focus on these four techniques because they represent major steps in PAPER's power spectrum pipeline, with several of them also being standard steps in general $21$ cm analyses.
\begin{itemize}
\item \textbf{Fringe-rate filtering:} Fringe-rate filtering is an averaging scheme for time-ordered data (\citealt{parsons_et_al2016}), and we explain the trade-offs of filtering in more detail in Section \ref{sec:SiglossOverview}. Broadly, a fringe-rate filter increases the sensitivity of a dataset and reduces the number of independent samples by an amount dependent on the width of the averaging window. However, it can also affect the presence of foregrounds and systematics. 
\item \textbf{Weighting:} A dataset can be weighted to emphasize certain spectral features and minimize others. One particular flavor of weighting is inverse covariance weighting, which is a generalized version of inverse variance weighting that also takes into account data correlations. This weighting has the effect of down-weighting correlated information (i.e. foregrounds) and up-weighting noise-like information (i.e. EoR). However, a challenge of inverse covariance weighting is in accurately defining a covariance matrix that best describes the data. 
\item \textbf{Bootstrapping:} Because theoretical models for covariance matrices are questionable and theoretical error estimation methods rely on assumptions that can be violated by data, applying theoretical errors to a $21$ cm measurement is often avoided. Hence, bootstrapping is a useful method for estimating errors of a dataset from itself. By randomly drawing many samples of the data, we obtain a sense of its inherent variance, though there are subtleties to consider such as the independence of values in a dataset.
\item \textbf{Jackknife testing:} A resampling technique useful for estimating bias, jackknives can be taken along different dimensions of a dataset to cross-validate results. In particular, null tests can be used to verify whether results are free of systematics (\citealt{keating_et_al2016}). An EoR detection must pass jack-knife and null tests.
\end{itemize}

@@

In this section we illustrate how signal loss arises in analysis. Signal loss refers to attenuation of the target cosmological signal in a power spectrum estimate. It can arise in a variety of ways in the analysis pipeline, such as fitting a polynomial during spectral calibration, applying a delay-domain filter, or by weighting data by itself. Here we focus on signal loss associated with applying a weighting matrix to data, a loss that can be significant depending on the choice of weighting and one that was previously underestimated in the PAPER-64 analysis.

@@

We begin with our data vector, $\textbf{x}$ which contains our measured visibilities in Jy. It has length ($N_{t}, N_{f}$), where $N_{t}$ is the number of time integrations and $N_{f}$ is the number of frequency channels. Visibilities are measurements of the Fourier transform of the sky along $2$ spatial dimensions, and since we are interested in $3$-dimensional Fourier-modes we only need to take one Fourier transform of our data along the line-of-sight dimension. We do this when forming the un-normalized power spectrum estimate $\hat{q}_{\alpha}$:

\begin{equation}
\label{eq:qhat}
\hat{q}_{\alpha} = \frac{1}{2}\textbf{x}^{\dagger}\textbf{R}\textbf{Q}^{\alpha}\textbf{R}\textbf{x}
\end{equation}

@@

\noindent where $\hat{\textbf{p}}$ is the normalized estimate of the true power spectrum. The data analyst has a choice for $\textbf{M}$. For simplicity in this section we set $\textbf{M} \propto \textbf{I}$, although we explore other cases for the analysis of PAPER-64 data as explained in Section \ref{sec:Bias}

In the next sections, we investigate the effects of weighting matrices on signal loss by experimenting with different matrices $\textbf{R}$ and examine their impact on the resulting power spectra estimate $\hat{\textbf{p}}$. Our ultimate goal in experimenting with weighting is to suppress foregrounds and investigate EoR losses associated with it.

@@

To build our intuition into how a particular choice of weighting, namely inverse covariance weighting, can lead to signal loss, we use a toy model. We construct a simple dataset that contains 2-dimensional data (representing visibility data with $100$ time integrations and $20$ frequency channels). This model represents realistic dimensions of about an hour of PAPER data which might be used for a power spectrum analysis. For PAPER-64 we use $\sim$$8$ hours of data for our analysis, but here we scale it down for this toy model with no loss of generality. We create a mock bright foreground signal, $\textbf{x}_{FG}$, as a complex sinusoid with a fixed frequency and a smoothly varying offset in time, a simplistic but realistic representation of a single bright source. We also create a mock EoR signal, $\textbf{x}_{EoR}$, as a complex, Gaussian random signal \cc{need to motivate why, since real EoR does have correlations}. Our combined data vector is $\textbf{x} = \textbf{x}_{FG} + \textbf{x}_{EoR}$, to which we apply different weighting schemes. The three data vectors are shown in Figure \ref{fig:toy_sigloss1}.

\begin{figure}
	\centering
	\includegraphics[trim={1.5cm 0.2cm 3.5cm 0.3cm},clip,width=\columnwidth]{plots/toy_sigloss1.png}
	\caption{Our toy model dataset contains a sinusoid foreground that varies in time and frequency and a random Gaussian signal as a mock EoR signal. Real parts are shown here.}
	\label{fig:toy_sigloss1}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[trim={1.5cm 0.2cm 3.5cm 0.3cm},clip,width=\columnwidth]{plots/toy_sigloss12.png}
	\includegraphics[trim={1.5cm 0.2cm 3.5cm 0.3cm},clip,width=\columnwidth]{plots/toy_sigloss13.png}
	\caption{The covariance matrices (top row) and inverse covariance-weighted data (bottom row) for FG only (left), EoR only (middle), and FG + EoR (right). Real parts are shown here.}
	\label{fig:toy_sigloss12}
\end{figure}

@@

It turns out that because we estimated $\hat{\textbf{C}}$ from our data, its eigenspectrum differs from the eigenspectrum of our true $\textbf{C}$, and this difference has consequences on our result. An eigenspectrum ranks the eigenvalues of a matrix from highest to lowest and can be thought of as a spectrum of weights that are given to each spectral mode in the data. In other words, the eigenvalues encode the strength of different shapes in the dataset. The eigenspectrum of the identity matrix $\textbf{I}$ is flat (all $1$'s) because it gives equal weighting to all modes. This is usually not the case for a covariance matrix, in which a sloped eigenspectrum means that modes are given different weights. When weighting data by an inverse covariance matrix, the modes with the highest eigenvalues are down-weighted the most.

\begin{figure}
	\centering
	\includegraphics[trim={0.3cm 0.2cm 0.3cm 0.3cm},clip,width=\columnwidth]{plots/toy_sigloss2.png}
	\caption{Eigenspectrum of $\hat{\textbf{C}}_{FG}$ (blue), $\hat{\textbf{C}}_{EoR}$ (red), and $\hat{\textbf{C}}_{FG+EoR}$ (green). The eigenspectrum of $\hat{\textbf{C}}_{FG}$ peaks at the zeroth eigenmode, due to the presence of only one sinusoid. These empirically estimated covariance matrices yield eigenspectra that are different from that of a true $\textbf{C}$, meaning that some eigenmodes (shapes in the data) may be down-weighted more significantly than they should be, producing signal loss.}
	\label{fig:toy_sigloss2}
\end{figure}

@@

In general, the strongest modes of $\hat{\textbf{C}}$ (highest eigenvalues) are more trustworthy than the weakest modes, since that's where we expect bright foregrounds --- the most prominent shapes --- to be. If bright foregrounds dominate, these highest eigenvalue modes therefore exist in both $\textbf{C}$ and $\hat{\textbf{C}}$. This is demonstrated in the toy model by the successful suppression of the foreground mode, where our estimated covariance matrix identifies the sinusoid and assigns it the highest eigenvalue (the peak in Figure \ref{fig:toy_sigloss2}).

@@

\begin{figure}
	\centering
	\includegraphics[trim={0.3cm 0.2cm 0.3cm 0.3cm},clip,width=\columnwidth]{plots/toy_sigloss7.png}
	\caption{Resulting power spectrum estimate for the fringe-rate filtered toy model simulation --- foregrounds only (blue), EoR only (red), and the weighted FG + EoR dataset (green). We use inverse covariance weighting where $\textbf{C}$ is derived from the data. There is a larger amount of signal loss than for the non-averaged data, a consequence of a steepened, weakly characterized eigenspectrum from using fewer independent modes in the data. The power spectrum estimate is also noisier, since most of the information used in constructing the estimate came from only a few eigenmodes.}
	\label{fig:toy_sigloss7}
\end{figure}

@@

As a first test, we know that our simulated EoR should have a covariance matrix that mimics the identity matrix, with its variance encoded along the diagonal. If we model $\textbf{C}_{EoR}$ as such, instead of computing it based on the dataset itself, and add it to $\hat{\textbf{C}}_{FG}$ (defined as $\hat{\textbf{C}}_{FG} = \langle\textbf{x}_{FG}\textbf{x}_{FG}^{\dagger}\rangle$) to obtain a final $\hat{\textbf{C}}$ to use in weighting, we see that there is negligible signal loss (Figure \ref{fig:toy_sigloss8}, upper left). This is because by modeling $\textbf{C}_{EoR}$, we avoid over-fitting small fluctuations in the data that our model doesn't know about (but an empirically derived $\hat{\textbf{C}}$ would). This is evident when comparing the green and red curves in Figure \ref{fig:toy_sigloss14}. We don't see perfect EoR recovery though, as the green and red curves differ at one $k$ value in Figure \ref{fig:toy_sigloss8}. This deviation is a consequence of the difference between the true and modeled covariance matrices.

@@

\begin{figure}
	\centering
	\includegraphics[trim={0.3cm 0.3cm 0.3cm 0.3cm},clip,width=\columnwidth]{plots/toy_sigloss14.png}
	\caption{We compare the eigenspectrum of an empirically calculated $\hat{\textbf{C}}$ (green) to that of four alternate weighting options, including modeling the covariance matrix of EoR (red), regularizing $\hat{\textbf{C}}$ by adding an identity matrix to it (blue), using only the first three eigenmodes of $\hat{\textbf{C}}$ (yellow), and multiplying an identity matrix with $\textbf{C}$ (magenta). }
	\label{fig:toy_sigloss14}
\end{figure}

The second panel (top right) in Figure \ref{fig:toy_sigloss8} uses a regularization method of setting $\hat{\textbf{C}} = \hat{\textbf{C}}_{ij} + \gamma\textbf{I}_{ij}$, where $i$$=$$j$$=$$20$ (number of frequencies) and $\gamma = 5$ (an arbitrary strength of $\textbf{I} for the purpose of this toy model$). By adding the identity matrix, element-wise, we are weighting the diagonal elements of the matrix more heavily than those off-diagonal, thereby flattening out its eigenspectrum. If all modes are given similar weights, we avoid down-weighting artificial modes more than others.

@@

Because foreground spectra are smooth in frequency, they preferentially show up at low delay, or $k$ modes. For a particular baseline length, there is a maximum delay imposed on foregrounds, which corresponds to the light-crossing time between the two antennas in the baseline. For longer baselines, this value increases, producing what is known as ``the wedge" (\citealt{parsons_et_al2012b}; \citealt{liu_et_al2014a}; \citealt{liu_et_al2014b}; \citealt{vedantham_et_al2012}; \citealt{thyagarajan_et_al2013}; \citealt{pober_et_al2013}; \citealt{datta_et_al2010}). The wedge describes a region in $k$-space contaminated by foregrounds, bounded by baseline length (which is proportional to $k_{\perp}$) and delay (which is proportional to $k_{\parallel}$). Properties of the wedge can be used to isolate and remove foregrounds, as done by \citet{ali_et_al2015}, \citet{parsons_et_al2014}, and \citet{jacobs_et_al2015}.

@@

There are analysis techniques to mitigate the effects of foreground leakage and prevent information from low $k's$ from spreading to high $k$ values. For example, narrow window functions can be used to minimize the covariance of a particular $k$ value with other ones (\citealt{liu_et_al2014b}). In other words, one can construct an estimator using OQE that forces a window function to have a minimum response to low $k$ values. The window function used in \citet{ali_et_al2015} is constructed in such a way, specifically to prevent foregrounds that live at low $k's$ from contaminating higher $k$-modes (see Section \ref{sec:Bias}). Minimizing foreground leakage in this way however, comes with the trade-off of compromising power spectrum sensitivity, since narrow window functions increases errors for each $k$-mode (\citealt{liu_et_al2014b}).

@@

We now present a case study of these same three themes using data from the PAPER experiment. We use the intuition developed through the toy model simulations in order to make a new analysis of the PAPER-64 dataset originally presented in \citet{ali_et_al2015} and obtain a revised power spectrum estimate.

As a brief review, PAPER is a dedicated $21$ cm experiment located in the Karoo Desert in South Africa. The PAPER-64 configuration consists of 64 dual-polarization drift-scan elements that are arranged in a grid layout. For our case study, we focus solely on Stokes I data from PAPER's $30$ m East/West baselines. For information about the backend system of PAPER-64, its observations, and data reduction pipeline, we refer the reader to \citet{parsons_et_al2010} and \citet{ali_et_al2015}.

The previously best published $21$ cm upper limit result from \citet{ali_et_al2015} uses $124$ nights of data to place a $2\sigma$ upper limit on $\Delta^{2}(k)$, defined as

@@

For our analysis, we use $8$ hours of LST (RA $0.5$-$8.6$ hours) and $51$ total baselines (\citet{ali_et_al2015} uses an RA range of $0$-$8.6$ hours). All power spectrum results are produced for a center frequency of $151$ MHz using a width of $10$ MHz ($20$ channels), identical to the analysis in \citet{ali_et_al2015}. We note that the PAPER-64 dataset that we use in this case study differs from that in \citet{ali_et_al2015} mainly by the fringe-rate filter. \cc{Danny says the next part is wrong, but that's what's in Zaki's paper... } In \citet{ali_et_al2015}, the applied filter was degraded by widening it in fringe-rate space. This was chosen in order to increase the number of independent modes and reduce signal loss. With the development of a robust method for assessing signal loss, we feel comfortable using a narrow filter --- the optimal fringe-rate filter --- in order to maximize sensitivity. This filter is computed for a fiducial $30$ m baseline at $150$ MHz, the center frequency in our band. The filter in both the fringe-rate domain and time domain is shown in Figure \ref{fig:frp}.

>>>>

\author{
Carina Cheng\altaffilmark{1},
Aaron R. Parsons\altaffilmark{1,2},
Matthew Kolopanis \altaffilmark{3},
Daniel C. Jacobs\altaffilmark{3}, 
Saul A. Kohn\altaffilmark{4},
Adrian Liu\altaffilmark{1,5}, 
James E. Aguirre\altaffilmark{4},
Zaki S. Ali\altaffilmark{1}, 
Richard F. Bradley\altaffilmark{6,7,8},
Gianni Bernardi\altaffilmark{9,10,11}, 
Chris L. Carilli\altaffilmark{12,13},
David R. DeBoer\altaffilmark{2}, 
Matthew R. Dexter\altaffilmark{2},
Joshua S. Dillon\altaffilmark{1},
Pat Klima\altaffilmark{7},
David H. E. MacMahon\altaffilmark{2},
David F. Moore\altaffilmark{4},
Chuneeta D. Nunhokee\altaffilmark{10},
Jonathan C. Pober\altaffilmark{14}, 
William P. Walbrugh\altaffilmark{9},
Andre Walker\altaffilmark{9}
}

\altaffiltext{1}{Astronomy Dept., U. California, Berkeley, CA}
\altaffiltext{2}{Radio Astronomy Lab., U. California, Berkeley CA}
\altaffiltext{3}{School of Earth and Space Exploration, Arizona State U., Tempe AZ}
\altaffiltext{4}{Dept. of Physics and Astronomy, U. Penn., Philadelphia PA} 
\altaffiltext{5}{Berkeley Center for Cosmological Physics, Berkeley, CA} 
\altaffiltext{6}{Dept. of Electrical and Computer Engineering, U. Virginia, Charlottesville VA}
\altaffiltext{7}{National Radio Astronomy Obs., Charlottesville VA}
\altaffiltext{8}{Dept. of Astronomy, U. Virginia, Charlottesville VA}
\altaffiltext{9}{Square Kilometer Array, S. Africa, Cape Town South Africa}
\altaffiltext{10}{Dept. of Physics and Electronics, Rhodes University, South Africa}
\altaffiltext{11}{Harvard-Smithsonian Cen. for Astrophysics, Cambridge MA}
\altaffiltext{12}{National Radio Astronomy Obs., Socorro NM}
\altaffiltext{13}{Cavendish Lab., Cambridge UK}
\altaffiltext{14}{Dept. of Physics, Brown University, Providence RI}

@@

\begin{abstract}
The Epoch of Reionization (EoR) is an uncharted era in our Universe's history during which the birth of the first stars and galaxies led to the ionization of neutral hydrogen in the intergalactic medium. There are many experiments investigating the EoR by tracing the $21$ cm line of neutral hydrogen, a signal which is very faint and difficult to isolate. With a new generation of instruments and a statistical power spectrum detection in our foreseeable future, it has become increasingly important to develop techniques that help maximize sensitivity and validate results. Additionally, it is imperative to understand the trade-offs between different methods and their effects on common power spectrum themes. In this paper, we focus on three major themes --- signal loss, power spectrum error bar estimation, and bias in measurements. We describe techniques that affect these themes using both toy models and data taken by the 64-element configuration of the Donald C. Backer Precision Array for Probing the Epoch of Reionization (PAPER). In particular, we highlight how detailed investigations of these themes have led to a revised, higher $21$ cm power spectrum upper limit from PAPER-64. This revised result mostly stems from an updated signal loss calculation for loss associated with empirically estimated covariances and supersedes results from previously published PAPER analyses.
\end{abstract}

@@

This paper is organized into two main sections. In Section \ref{sec:Themes} we introduce the three themes of our focus, using a toy model to develop intuition for each one. In Section \ref{sec:CaseStudy} we present a case study into each theme using data from the PAPER-64 array, highlighting key changes from \citet{ali_et_al2015}, henceforth known as P64, which have led to a revised PAPER-64 power spectrum result (\cc{cite retraction or kolopanis paper}). We conclude in Section \ref{sec:Con}.

@@

The following techniques each have advantages when it comes to maximizing sensitivity and understanding systematics in data. However, some have limitations, and we will discuss circumstances in which there are trade-offs. We choose to focus on these four techniques because they represent major steps in PAPER's power spectrum pipeline, with several of them also being standard steps in general $21$ cm analyses.
\begin{itemize}
\item \textbf{Fringe-rate filtering:} Fringe-rate filtering is an averaging scheme for time-ordered data (\citealt{parsons_et_al2016}). Broadly, a fringe-rate filter increases the sensitivity of a dataset and reduces the number of independent samples by an amount dependent on the width of the averaging window. However, it can also affect the presence of foregrounds and systematics. We explain the trade-offs of filtering in more detail in Section \ref{sec:SiglossOverview}.
\item \textbf{Weighting:} A dataset can be weighted to emphasize certain spectral features and minimize others. One particular flavor of weighting is inverse covariance weighting, which is a generalized version of inverse variance weighting that also takes into account data correlations. This weighting has the effect of down-weighting correlated information (i.e. foregrounds) and up-weighting noise-like information (i.e. EoR). However, a challenge of inverse covariance weighting is in accurately defining a covariance matrix that best describes the data. 
\item \textbf{Bootstrapping:} Because theoretical models for covariance matrices are questionable and theoretical error estimation methods rely on assumptions that can be violated by data, applying theoretical errors to a $21$ cm measurement is often avoided. Hence, bootstrapping is a useful method for estimating errors of a dataset from itself. By randomly drawing many samples of the data, we obtain a sense of its inherent variance, though there are subtleties to consider such as the independence of values in a dataset.
\item \textbf{Jackknife testing:} A resampling technique useful for estimating bias, jackknives can be taken along different dimensions of a dataset to cross-validate results. In particular, null tests can be used to verify whether results are free of systematics (\citealt{keating_et_al2016}). An EoR detection must pass jack-knife and null tests.
\end{itemize}

@@

In this section we illustrate how signal loss arises in analysis. Signal loss refers to attenuation of the target cosmological signal in a power spectrum estimate. It can arise in a variety of ways in the analysis pipeline, such as fitting a polynomial during spectral calibration, applying a delay-domain filter, or by weighting data by itself. Here we focus on signal loss associated with applying a weighting matrix to data, a loss that can be significant depending on the choice of weighting and one that was previously underestimated in the P64 analysis

@@

We begin with our data vector, $\textbf{x}$, which contains our measured visibilities in Jy. It has length ($N_{t}, N_{f}$), where $N_{t}$ is the number of time integrations and $N_{f}$ is the number of frequency channels. Visibilities are measurements of the Fourier transform of the sky along $2$ spatial dimensions, and since we are interested in $3$-dimensional Fourier-modes we only need to take one Fourier transform of our data along the line-of-sight dimension. We do this when forming the un-normalized power spectrum estimate $\hat{q}_{\alpha}$:

\begin{equation}
\label{eq:qhat}
\hat{q}_{\alpha} = \frac{1}{2}\textbf{x}^{\dagger}\textbf{R}\textbf{Q}^{\alpha}\textbf{R}\textbf{x}.
\end{equation}

@@

\noindent where $\hat{\textbf{p}}$ is the normalized estimate of the true power spectrum. The data analyst has a choice for $\textbf{M}$ --- for simplicity in this section we set $\textbf{M} \propto \textbf{I}$, although we explore other cases for the analysis of PAPER-64 data as explained in Section \ref{sec:Bias}

In the next three sections, we investigate the effects of weighting matrices on signal loss by experimenting with different matrices $\textbf{R}$ and examine their impact on the resulting power spectra estimate $\hat{\textbf{p}}$. Our ultimate goal in experimenting with weighting is to suppress foregrounds and investigate EoR losses associated with it.

@@

To build our intuition into how a particular choice of weighting, namely inverse covariance weighting, can lead to signal loss, we use a toy model. We construct a simple dataset that contains 2-dimensional data (representing visibility data with $100$ time integrations and $20$ frequency channels). This model represents realistic dimensions of about an hour of PAPER data which might be used for a power spectrum analysis. For PAPER-64 we use $\sim$$8$ hours of data for our analysis, but here we scale it down for this toy model with no loss of generality. We create a mock bright foreground signal, $\textbf{x}_{FG}$, as a complex sinusoid with a fixed frequency and a smoothly varying offset in time, a simplistic but realistic representation of a single bright source. We also create a mock EoR signal, $\textbf{x}_{EoR}$, as a complex, Gaussian random signal. A more realistic EoR signal would have a sloped power spectrum in $\hat{\textbf{p}}$(k) (instead of flat, as in the case of white noise), which could be simulated by introducing frequency correlations into the mock EoR signal. However, in this paper we treat all $k$'s separately, so a simplistic white noise approximation can be used. Our combined data vector is then $\textbf{x} = \textbf{x}_{FG} + \textbf{x}_{EoR}$, to which we apply different weighting schemes. The three data vectors are shown in Figure \ref{fig:toy_sigloss1}.

\begin{figure}
	\centering
	\includegraphics[trim={1.5cm 0.2cm 3.5cm 0.3cm},clip,width=\columnwidth]{plots/toy_sigloss1.png}
	\caption{Our toy model dataset contains a sinusoid foreground that varies smoothly in time and a random Gaussian signal as a mock EoR signal. Real parts are shown here.}
	\label{fig:toy_sigloss1}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[trim={1.5cm 0.1cm 3.5cm 0.1cm},clip,width=\columnwidth]{plots/toy_sigloss12.png}
	\includegraphics[trim={1.5cm 0.1cm 3.5cm 0.1cm},clip,width=\columnwidth]{plots/toy_sigloss13.png}
	\caption{The covariance matrices (top row) and inverse covariance-weighted data (bottom row) for FG only (left), EoR only (middle), and FG + EoR (right). Real parts are shown here.}
	\label{fig:toy_sigloss12}
\end{figure}

@@

It turns out that because we estimated $\hat{\textbf{C}}$ from our data, its eigenspectrum differs from the eigenspectrum of our true $\textbf{C}$, and this difference has consequences on our result. An eigenspectrum ranks the eigenvalues of a matrix from highest to lowest and can be thought of as a spectrum of weights that are given to each spectral mode in the data. In other words, the eigenvalues encode the strength of different shapes in the dataset. The eigenspectrum of the identity matrix $\textbf{I}$ is flat (all $1$'s) because it gives equal weighting to all modes. This is usually not the case for a covariance matrix, for which a sloped eigenspectrum means that modes are given different weights. The modes with the highest eigenvalues are down-weighted the most.

\begin{figure}
	\centering
	\includegraphics[trim={0.3cm 0.2cm 0.3cm 0.3cm},clip,width=\columnwidth]{plots/toy_sigloss2.png}
	\caption{Eigenspectrum of $\hat{\textbf{C}}_{FG}$ (blue), $\hat{\textbf{C}}_{EoR}$ (red), and $\hat{\textbf{C}}_{FG+EoR}$ (green). The eigenspectrum of $\hat{\textbf{C}}_{FG}$ peaks at the zeroth eigenmode, due to the presence of only one sinusoid. These empirically estimated covariance matrices have eigenspectra that are different from that of a true $\textbf{C}$, meaning that some eigenmodes (shapes in the data) may be down-weighted more significantly than they should be, producing signal loss.}
	\label{fig:toy_sigloss2}
\end{figure}

@@

In general, the strongest modes of $\hat{\textbf{C}}$ (highest eigenvalues) are more trustworthy than the weakest modes, since that's where we expect bright foregrounds --- the most prominent shapes --- to be. If bright foregrounds dominate, these highest eigenvalue modes would exist in both $\textbf{C}$ and $\hat{\textbf{C}}$. This is demonstrated in the toy model by the successful suppression of the foreground mode, where our estimated covariance matrix identifies the sinusoid shape and assigns it the highest eigenvalue (the peak in Figure \ref{fig:toy_sigloss2}).

@@

\begin{figure}
	\centering
	\includegraphics[trim={0.3cm 0.2cm 0.3cm 0.3cm},clip,width=\columnwidth]{plots/toy_sigloss7.png}
	\caption{Resulting power spectrum estimate for the fringe-rate filtered toy model simulation --- foregrounds only (blue), EoR only (red), and the weighted FG + EoR dataset (green). We use inverse covariance weighting where $\textbf{C}$ is computed from the data. There is a larger amount of signal loss than for the non-averaged data, a consequence of a steepened, weakly characterized eigenspectrum from using fewer independent modes in the data. The power spectrum estimate is also noisier, since most of the information used in constructing the estimate came from only a few eigenmodes.}
	\label{fig:toy_sigloss7}
\end{figure}

@@

As a first test, we know that our simulated EoR signal should have a covariance matrix that mimics the identity matrix, with its variance encoded along the diagonal. If we model $\textbf{C}_{EoR}$ as such, instead of computing it based on the dataset itself, and add it to $\hat{\textbf{C}}_{FG}$ (defined as $\hat{\textbf{C}}_{FG} = \langle\textbf{x}_{FG}\textbf{x}_{FG}^{\dagger}\rangle$) to obtain a final $\hat{\textbf{C}}$ to use in weighting, we see that there is negligible signal loss (Figure \ref{fig:toy_sigloss8}, upper left). This is because by modeling $\textbf{C}_{EoR}$, we avoid over-fitting small fluctuations in the data that our model doesn't know about (but an empirically derived $\hat{\textbf{C}}$ would). This is evident when comparing the green and red curves in Figure \ref{fig:toy_sigloss14}. We don't see perfect EoR recovery though, as the green and red curves vary slightly in Figure \ref{fig:toy_sigloss8}. This deviation is a consequence of the difference between the true and modeled covariance matrices.

@@

\begin{figure}
	\centering
	\includegraphics[trim={0.3cm 0.3cm 0.3cm 0.3cm},clip,width=\columnwidth]{plots/toy_sigloss14.png}
	\caption{We compare the eigenspectrum of an empirically calculated $\hat{\textbf{C}}$ (green) to that of four alternate weighting options, including modeling the covariance matrix of EoR (red), regularizing $\hat{\textbf{C}}$ by adding an identity matrix to it (blue), using only the first three eigenmodes of $\hat{\textbf{C}}$ (yellow), and multiplying an identity matrix with $\hat{\textbf{C}}$ (magenta). All eigenspectra (except the green) are relatively flat and don't exhibit signal loss.}
	\label{fig:toy_sigloss14}
\end{figure}

The second panel (top right) in Figure \ref{fig:toy_sigloss8} uses a regularization method of setting $\hat{\textbf{C}} = \hat{\textbf{C}}_{ij} + \gamma\textbf{I}_{ij}$, where $i$$=$$j$$=$$20$ (number of frequencies) and $\gamma = 5$ (an arbitrary strength of $\textbf{I}$ for the purpose of this toy model). By adding the identity matrix, element-wise, we are weighting the diagonal elements of the matrix more heavily than those off-diagonal, thereby flattening out its eigenspectrum. If all modes are given similar weights, we avoid down-weighting artificial modes more than others.

@@

Because foreground spectra are smooth in frequency, they preferentially show up at low delay, or $k$ modes. For a particular baseline length, there is a maximum delay imposed on foregrounds, which corresponds to the light-crossing time between the two antennas in the baseline. For longer baselines, this value increases, producing what is known as ``the wedge" (\citealt{parsons_et_al2012b}; \citealt{liu_et_al2014a}; \citealt{liu_et_al2014b}; \citealt{vedantham_et_al2012}; \citealt{thyagarajan_et_al2013}; \citealt{pober_et_al2013}; \citealt{datta_et_al2010}). The wedge describes a region in $k$-space contaminated by foregrounds, bounded by baseline length (which is proportional to $k_{\perp}$) and delay (which is proportional to $k_{\parallel}$). Properties of the wedge can be used to isolate and remove foregrounds, as done by P64, \citet{parsons_et_al2014}, and \citet{jacobs_et_al2015}.

@@

There are analysis techniques to mitigate the effects of foreground leakage and prevent information from low $k's$ from spreading to high $k$ values. For example, narrow window functions can be used to minimize the covariance of a particular $k$ value with other ones (\citealt{liu_et_al2014b}). In other words, one can construct an estimator using OQE that forces a window function to have a minimum response to low $k$ values. The window function used in P64 is constructed in such a way, specifically to prevent foregrounds that live at low $k's$ from contaminating higher $k$-modes (see Section \ref{sec:Bias}). Minimizing foreground leakage in this way however, comes with the trade-off of compromising power spectrum sensitivity, since narrow window functions increases errors for each $k$-mode (\citealt{liu_et_al2014b}).

@@

We now present a case study of these same three themes using data from the PAPER experiment. We use the intuition developed through the toy model simulations in order to make a new analysis of the PAPER-64 dataset originally presented in P64 and obtain a revised power spectrum estimate.

As a brief review, PAPER is a dedicated $21$ cm experiment located in the Karoo Desert in South Africa. The PAPER-64 configuration consists of 64 dual-polarization drift-scan elements that are arranged in a grid layout. For our case study, we focus solely on Stokes I data from PAPER's $30$ m East/West baselines. For information about the backend system of PAPER-64, its observations, and data reduction pipeline, we refer the reader to \citet{parsons_et_al2010} and P64.

The previously best published $21$ cm upper limit result from P64 uses $124$ nights of data to place a $2\sigma$ upper limit on $\Delta^{2}(k)$, defined as

@@

For our analysis, we use $8$ hours of LST (RA $0.5$-$8.6$ hours) and $51$ total baselines (P64 uses an RA range of $0$-$8.6$ hours). All power spectrum results are produced for a center frequency of $151$ MHz using a width of $10$ MHz ($20$ channels), identical to the analysis in P64. We note that the PAPER-64 dataset that we use in this case study differs from that in P64 mainly by the fringe-rate filter. \cc{Danny says the next part is wrong, but that's what's in Zaki's paper... } In P64, the applied filter was degraded by widening it in fringe-rate space. This was chosen in order to increase the number of independent modes and reduce signal loss. With the development of a robust method for assessing signal loss, we feel comfortable using a narrow filter --- the optimal fringe-rate filter --- in order to maximize sensitivity. This filter is computed for a fiducial $30$ m baseline at $150$ MHz, the center frequency in our band. The filter in both the fringe-rate domain and time domain is shown in Figure \ref{fig:frp}.
